\vspace{1.0cm} \Large \bf Integrating influenza antigenic dynamics with molecular evolution

### 2013-05-14 response.tex

\textsl{* The authors say, ``If a single map is desired, then the MDS method may be preferred while the BMDS method provides flexibility, allowing other sources of information, such as genetic data, to be more easily incorporated.'' Is this the first part of this sentence true? The BMDS method introduced in this paper is such a clear improvement on the original method in the 2004 Smith et al Science paper that this new approach should always be preferred. I don't see any difficulties with summarizing the posterior of the BMDS as done in this paper. It might be possible to make some non-Bayesian (such as maximum-likelihood version) of the method in this paper that included genetic data / dates / virus+serum effects, and then argue that such a maximum-likelihood version could be preferred to the BMDS method here. But if the comparison is simply with the 2004 Smith et al Science approach (which only uses HI data without dates / phylogeny / virus+serum effects), then the current approach is clearly better due to the inclusion of additional data, regardless of one's philosophical preference for Bayesian versus maximum likelihood.}

\textsl{* In the text, the authors introduce the three beta coefficients without first defining what they mean in either intuitive or mathematical terms.}

\textsl{* The authors say, ``we find that interference between sister clades, e.g. B/Vic and B/Yam, appears stronger than interference between more divergent clades, e.g. B/Vic and A/H3N2.'' However, this statement is not linked to any analysis/data - where is the figure or table?}

This statement comes from the relative magnitudes of the $\beta_s$, representing the effect of antigenic drift on sister clades, and $\beta_t$, representing the effect of drift on more divergent clades.  We find $\beta_s = -0.61$ and $\beta_t = -0.32$.  We have clarified the text accordingly.

@@

\textsl{* It would be helpful to have phylogenetic trees for the four clades as supporting figures, since as best I understand Figure 2 and 3 do not represent the phylogenetic relationships, but only show the influence of those relationships on the antigenic map. Also, are the phylogenetic trees built on just HA1 or the whole HA protein?}

\textsl{* Although this is a stylistic rather than a substantive comment, the paper ends with a weak conclusion. The authors present a lot of complex results that could use some intuitive interpretations. Instead, their second-to-last sentence is a technical point about the phylogenetic methodology.}

\textsl{* The authors should be commended on writing a clear Methods section, particularly as it relates to the antigenic cartography.}

Thank you. We attempted to make the cartography methods as clear as possible.

@@

\textsl{* Page 16, there is an extraneous ``the'' before the word ``observing.''}

Fixed.

\textsl{* In equation 12, what is meant by the x subscript given to the sigma?}

The $\sigma_x$ in the diffusion process is describing variance among virus locations $\mathbf{X}$.  The $x$ subscript is meant to differentiate this from serum locations $\mathbf{Y}$.  This has been clarified in the text.

### 2013-05-14 flux.tex
Fixes #15 - Strength of sister clade interference.

These findings clearly demonstrate the epidemiological importance of antigenic drift and show that incidence across  A/H3N2, A/H1N1, B/Vic and B/Yam is strongly influenced by levels of antigenic drift within each clade, consistent with antigenic drift increasing the susceptible pool available to a clade.
The finding of a negative correlation between incidence within one clade and antigenic drift in other clades supports the hypothesis that epidemiological interference exists between clades \cite{Ferguson03, Goldstein11}.
Interestingly, we find that interference between sister clades, e.g.\ B/Vic and B/Yam, appears stronger than interference between more divergent clades, e.g.\ B/Vic and A/H3N2.
If interference is mediated by CTL epitopes, then we would expect interference to decrease with evolutionary distance \cite{Voeten00}.

>>>>

These findings clearly demonstrate the epidemiological importance of antigenic drift and show that incidence across  A/H3N2, A/H1N1, B/Vic and B/Yam is strongly influenced by levels of antigenic drift within each clade, consistent with antigenic drift increasing the susceptible pool available to a clade.
The finding of a negative correlation between incidence within one clade and antigenic drift in other clades supports the hypothesis that epidemiological interference exists between clades \cite{Ferguson03, Goldstein11}.
Interestingly, we find that interference between sister clades, e.g.\ B/Vic and B/Yam, appears stronger than interference between more divergent clades, e.g.\ B/Vic and A/H3N2 (Table~\ref{interferencetable}, $\bsister = -0.61$ vs $\bother = -0.32$).
If interference is mediated by CTL epitopes, then we would expect interference to decrease with evolutionary distance \cite{Voeten00}.
